---
title: "Topf2018.Dataanalysis_FINALforPUBLI"
author: "Lukas Wille"
date: "1/2/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries I need, echo=FALSE, message=FALSE, error=FALSE}
library(dplyr)
library(tidyr)
library(purrr)
library(reshape2)
library(ggplot2)
library(gridExtra) # fuer grid.arrange
library(LambertW)
library(emmeans)
library(ARTool) # For Aligned Rans transformation ANOVA of Root Rot index
library(vegan)
library(ecoCopula) # Graphical model (Gordana Popovic)
library(ordinal) # for ecoCopula
library(mvabund) # analysis of multivariate abundance data
library(labdsv)    # Ordination analysis in community ecology for ecoCopula analysis
library(igraph)
library(mgcv)
#library(SpiecEasi)
```

```{r WD setup, echo=FALSE, error=TRUE}
setwd("~/ownCloud/FiblWulche/Dataanalysis/Topfversuch2018/FinalAnalysisforPubli") # MAC
setwd("C:/Users/lukas.wille/ownCloud/FiblWulche/Dataanalysis/Topfversuch2018/FinalAnalysisforPubli") # WIN
```

```{r Read-in ready Data, echo=FALSE, warning=FALSE, error=TRUE}
setwd("~/ownCloud/FiblWulche/Dataanalysis/Topfversuch2018/DataframeszumSchaffe")    # MAC
setwd("C:/Users/lukas.wille/ownCloud/FiblWulche/Dataanalysis/Topfversuch2018/DataframeszumSchaffe") # WIN

# Phenotypic data
Topf18.S <- readRDS("Topf18.S.rds")
Topf18.NS <- readRDS("Topf18.NS.rds")

Topf18.4.8 <- readRDS("Topf18.4.8.rds")     # only 4 soils, 8 Acc's, S and NS
Topf18.4.8.S <- readRDS("Topf18.4.8.S.rds")
Topf18.4.8.NS <- readRDS("Topf18.4.8.NS.rds")

Topf18.4.8.3sick <- readRDS("Topf18.4.8.3sick.rds")

# qPCR data for 10 fungis
Topf18.FUNGI <- readRDS("Topf18.FUNGI.rds")   # DFs with only qPCR Conc Data (ONE & TWO are the 2 tech reps.; Fungal name without suffix ONE/TWO ist the mean between the two...)
Topf18.FUNGI.3SICK <- readRDS("Topf18.FUNGI.3SICK.rds")

Topf18.FUNGI.S <- readRDS("Topf18.FUNGI.S.rds")
```

```{r helper vectors, echo=FALSE}
thefungis <- names(Topf18.FUNGI)[c(9,12,15,18,21,24,27,30,33,36)]  # Taxa names of the fungis
thefungis <- set_names(thefungis)    #Purrr Package for set_names function

theaccs <- set_names(levels(Topf18.NS$acc))

soil_names <- c(`F` = "Feldbach",`H` = "Kirchlindach",`L` = "Puch",`W` = "Neu-Eichenberg")

soil_nms <- c(`F` = "F",`H` = "K",`L` = "P",`W` = "N")

Kolor = c("#009900","#FF0033","#CC9900","#663333")  # colors for the 4 soils

fungKolor = c("#A1D99B", "#800026", "darkgreen", "#FC9272", "#3399FF", "#C994C7", "#E6550D", "black", "#FED976", "pink")
```


# Analysis of shoot dry weight (SDW) - S vs NS treatment

I use this analysis to show that the 3 sick soils are sick, and Feldbach healthy. This gives a Supplementary figure.
https://authorservices.wiley.com/asset/photos/electronic_artwork_guidelines.pdf
https://nph.onlinelibrary.wiley.com/hub/journal/14698137/about/author-guidelines

```{r Fig S1, echo=FALSE, warning=FALSE}
plotSDW <- ggplot(aes(treat, s.dw.1), data=Topf18.4.8) +
  geom_boxplot() +
  stat_summary(fun.y=mean, geom="point", shape=4, size=4) +
  facet_wrap(~soil, labeller = labeller(soil = soil_names)) +
  theme_bw() +
    theme(legend.title = element_blank(),
          panel.background =  element_blank(),
          panel.grid = element_blank()) +
  labs(y=expression("Shoot dry weight [g]"),
       x=expression("Soil treatment")) +
  ggpubr::stat_compare_means(method = "t.test", label.y = 0.6)   # to directly plot p-values of t.test into plot :) For New Phyt I should give the df's as well, however, I do not see that in all the published papers...

plotSDW
#ggsave("SDW_SvsNS_in4soils.eps", plotSDW, units="cm", width=18, height=18, dpi=300, device = "eps")    # I save only once. 80mm width is one col in New Phytologist; 180mm is two cols
```


# Analysis of relative shoot dry weight (SDWRel)

## Summarize SDWrel (N, mean +- SD)

```{r summary pheno for soils and accs, echo=FALSE, message=FALSE}

sumbysoilacc_SDWrel <- Topf18.4.8.NS %>% group_by(soil, acc) %>%
  summarise(N_not_NA = sum(!is.na(s.dw.ratioSNS)),
            mean= mean(s.dw.ratioSNS, na.rm = TRUE),
            sd = sd(s.dw.ratioSNS, na.rm = TRUE))

sumbysoilacc_SDWrel
```
-> C2 in F and W has two observations (all others have 3 at least). I leave C2 in for the analysis of SDWRel.



## Plot SDWrel: 4 soils over 8 genotypes & 4 soils-8 genotypes individually: Mean +- SE

```{r  beautiful SDWRel plot, echo=FALSE}

#Panel A: The boxplot of the 4 soils
gh <- ggplot() +
  geom_boxplot(data = Topf18.4.8.NS, aes(x = soil, y = s.dw.ratioSNS, fill = soil), outlier.size = 0.5) +
  stat_summary(data = Topf18.4.8.NS, aes(x = soil, y = s.dw.ratioSNS), fun=mean, geom="point", shape=4, size=2) +
  scale_fill_manual(values = Kolor) +
  theme_classic() +
  scale_y_continuous(sec.axis = dup_axis(breaks = 0),        # put a fake right y-axis as border...
                     expand = c(0, 0),
                     breaks = c(0,1,2)) +                     
  coord_cartesian(ylim = c(-0.13, 2.5)) +  # set explicit range
  geom_hline(yintercept = 2.5) +
  geom_hline(yintercept = 1, lty = 2, size = 0.5) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_text(size=8),
        axis.title.y.right = element_blank(),               # hide right axis title
        axis.text.y.right = element_blank(),
        axis.ticks.y.right = element_blank(),
        axis.title.y.left = element_text(size=8),
        legend.title = element_blank(),
        panel.background =  element_blank(),
        panel.grid = element_blank()) +
  labs(tag="A",
       y=expression(SDW[italic(Rel.)]),
       x=expression("Soil")) +
  annotate("text", x = 1, y = 1.8, label = "a",  size=3) +
  annotate("text", x = 2, y = 1.6, label = "bc", size=3) +
  annotate("text", x = 3, y = 1.5, label = "c",  size=3) +
  annotate("text", x = 4, y = 1.7, label = "b",  size=3)

#Panel B: The kindofainteractionplot showing the genotypes -> this plot also generates the legend for both panels
sum <- sumbysoilacc_SDWrel
sum <- sum %>% mutate(se=sd/sqrt(N_not_NA))   #add SE to the sum file
#sum$acc <- factor(sum$acc, c("resistant", "C1", "S91", "S134", "S64", "G78", "susceptible", "C2", "S22", "G89")) #the empty "fake" factors are placeholders for the legend before 18.1.21
sum$acc <- factor(sum$acc, c("resistant", "S91", "S134", "G78", "C1", "S64", "susceptible", "G89", "S22", "C2")) # from 18.1.21 on



gc <- ggplot(data = sum) +
geom_linerange(aes(x = acc, ymin = mean - se, ymax = mean + se, color = soil)) +
geom_point(aes(x = acc, y = mean, shape = acc, color = soil), size=1.5) +
facet_wrap(~soil, nrow = 1, labeller = labeller(soil = soil_names)) +
scale_shape_manual(values = c(NA, 19, 15, 18, 17, 20, NA, 2, 1, 0), drop=FALSE) +   # !
scale_color_manual(values = Kolor, labels = soil_names) +
scale_fill_manual(labels = soil_names) +
theme_classic() +
scale_y_continuous(labels = function(x){paste(x, "-")},    
                     sec.axis = dup_axis(breaks = 0),        
                     expand = c(0, 0),
                    breaks = c(0,1,2)) +                     # put a fake right y-axis as border...
coord_cartesian(ylim = c(-0.13, 2.5)) +  # set explicit range
geom_hline(yintercept = 2.5 ) +
geom_hline(yintercept = 1, lty = 2, size = 0.5) +
theme(
      axis.text.x = element_blank(),
      axis.title.x = element_text(size=8),
      axis.ticks.x = element_blank(),
      axis.title.y.right = element_blank(),               # hide right axis title
      axis.text.y = element_blank(),
      axis.ticks.y.right = element_blank(),
      panel.spacing = unit(0, "lines"),
      strip.background = element_blank(),
      strip.text.x = element_blank(),
      legend.title = element_blank(),
      panel.background =  element_blank(),
      panel.grid = element_blank()
      ) +
labs(tag="B",
        y=expression(""),
        x=expression("Genotype")) +
guides(#shape = guide_legend(ncol = 1, override.aes = list(shape = c(NA, 17, 19, 15, 20, 18, NA, 0, 1, 2))),
       shape = guide_legend(ncol = 1, override.aes = list(shape = c(NA, 19, 15, 18, 17, 20, NA, 2, 1, 0))),
  color = guide_legend(ncol = 1, override.aes = list(shape = c(rep(20,4)), size=2)))

#put the two panels A + B  & Save
#get legends (note the 4 soils legend is already in the big panel), and arrange them together

props <- ggpubr::ggarrange(gh + theme(legend.position="none"),
                           gc,
                           widths = c(1,4,3),
                           nrow=1)
#ggsave("SDWrelBoxplotwithInteractionplot_UNFERTIG.tiff", props, units="cm", width=18, height=6, dpi=300, compression="lzw")    # I save only once. "UNFERTIG" weil noch weiterbearbeitet in gimp
```



## Regression/ANOVA of SDWrel with factor 'acc'
-> Original SDWRel data is Lambert W x Fx transformed ("Gaussianized", Goerg 2015)

```{r Gaussianize SDWRel}
# reduce dataset to variable in focus and design variables
Topf18.4.8.NS.SDWRel.GAUSS <- Topf18.4.8.NS[,c(1:10,27)]

p <- Topf18.4.8.NS.SDWRel.GAUSS[,c(2,11)]  # only LABEL and SDWrel
p <- p[complete.cases(p), ]   # the Gaussianize function does not work with NAs; so I take them out, and put them back after the whole procedure

# do the transformation
G_List_SDWRel <- LambertW::Gaussianize(p[,2], type="hh", method = "MLE", return.tau.mat = TRUE)
p$SDWRel_G<-as.vector(G_List_SDWRel$input) 
Topf18.4.8.NS.SDWRel.GAUSS <- merge(Topf18.4.8.NS.SDWRel.GAUSS, p[c("LABEL", "SDWRel_G")], all.x=TRUE)

# Remove objects
rm(p)
```

After transformation I  do the linear Model:

```{r formulate the Model and get Anova 4 soils acc}
LM.48.SDW_G <- lm(SDWRel_G ~ soil * acc + rep, Topf18.4.8.NS.SDWRel.GAUSS,
                     contrasts=list(soil=contr.sum, acc=contr.sum, rep=contr.sum))   # the "contrasts" statement needs to be set like this for proper calculation of Anova type III
car::Anova(LM.48.SDW_G, type="III")

#Model diagnostic via residuals
par(mfrow=c(1,2))
scatter.smooth(fitted(LM.48.SDW_G), resid(LM.48.SDW_G)); abline(h=0,lty=2)
title("Tukey-Anscombe Plot")
car::qqPlot(resid(LM.48.SDW_G), pch=3, cex=0.5, id=list(n=5, cex=0.7, col="red"), line="quartiles")
```

Result: soil & acc are signf.; interaction not. Residuals look ok.

## Post-hoc analysis: Planned contrasts between resistant and susceptible genotypes for each soil

```{r posthoc SDW orthogonal reslevel within soil}

emm.LM.SoilxAcc.SDW_G <- emmeans(LM.48.SDW_G, c("acc", "soil"))
levels(Topf18.4.8.NS.SDWRel.GAUSS$acc)  # check the order of the accs: "C1" "C2" "G78" "G89" "S134" "S22" "S64" "S91" -> in order to specify the dummy variables for the contrasts correctly
Contrasts = list(Res_vs_susceptible = c(1/5, -1/3, 1/5, -1/3, 1/5, -1/3, 1/5, 1/5))  # I need to specify the contrasts so that the 5 resistant and the 3 susceptible genotypes sum-up to 0.
contrast(emm.LM.SoilxAcc.SDW_G, Contrasts, by="soil")

```

I see that R-S do not significantly differ in F. But S are significantly lower than R in the three infested soils.  
 

## Regression/ANOVA of SDWrel with factor 'reslevel'
-> to test differences between the two groups resistant and susceptible a new model is fit where the factor 'acc' is replaced by 'reslevel'  

```{r formulate the Model and get Anova 4 soils resistance}
LM.48.SDW_G_reslevel <- lm(SDWRel_G ~ soil * reslevel + rep, Topf18.4.8.NS.SDWRel.GAUSS,
                     contrasts=list(soil=contr.sum, reslevel=contr.sum, rep=contr.sum))   # the "contrasts" statement needs to be set like this for proper calculation of Anova type III
car::Anova(LM.48.SDW_G_reslevel, type="III")

#Model diagnostic via residuals
par(mfrow=c(1,2))
scatter.smooth(fitted(LM.48.SDW_G_reslevel), resid(LM.48.SDW_G_reslevel)); abline(h=0,lty=2)
title("Tukey-Anscombe Plot")
car::qqPlot(resid(LM.48.SDW_G_reslevel), pch=3, cex=0.5, id=list(n=5, cex=0.7, col="red"), line="quartiles")
```

-> soil, reslevel & interaction are significant. Residuals look ok.  
I do not show this result in the publication. However, the post-hoc analysis of reslevel within each soil is interesting:

```{r reslevel 4 soils post-hoc}

emm.LM.SoilxAcc.SDW_G_reslevel <- emmeans(LM.48.SDW_G_reslevel, c("soil", "reslevel"))
POSTHOC_SDW_G_TukeySoilxAcc_reslevel <- contrast(emm.LM.SoilxAcc.SDW_G_reslevel, method="pairwise", by="soil")
POSTHOC_SDW_G_TukeySoilxAcc_reslevel

```





## Regression/ANOVA of SDWrel only in the 3 infested soils with factor 'acc

```{r formulate the Model and get Anova 3 soils acc}
LM.48.SDW_G_3sick <- lm(SDWRel_G ~ soil * acc + rep, Topf18.4.8.NS.SDWRel.GAUSS[Topf18.4.8.NS.SDWRel.GAUSS$soil!="F",],
                     contrasts=list(soil=contr.sum, acc=contr.sum, rep=contr.sum))   # the "contrasts" statement needs to be set like this for proper calculation of Anova type III
car::Anova(LM.48.SDW_G_3sick, type="III")

#Model diagnostic via residuals
par(mfrow=c(1,2))
scatter.smooth(fitted(LM.48.SDW_G_3sick), resid(LM.48.SDW_G_3sick)); abline(h=0,lty=2)
title("Tukey-Anscombe Plot")
car::qqPlot(resid(LM.48.SDW_G_3sick), pch=3, cex=0.5, id=list(n=5, cex=0.7, col="red"), line="quartiles")
```
In order to compare the genotypic means in the 3 sick soils post-hoc analysis (Tukeys HSD) are done. Because there is no significant soil x acc interaction, estimated means are calculated for each genotype over the 3 sick soils (note the warning of the emmeans() function)

```{r post hoc analysis 3 sick soils acc}

#contrast(emmeans(LM.48.SDW_G_3sick, ~ acc), method="pairwise")
multcomp::cld(emmeans(LM.48.SDW_G_3sick, ~ acc), alpha=0.05, Letters=letters, adjust="tukey")  # letter display for the pairwise testing
```


# Analysis of qPCR data

## Summary of qPCR data (mean & SD for each genotype in each soil) -> Supp Tab 5 (for this table I used the full data including data on C2 in F & W soils)

```{r fungi stack, echo=FALSE, message=FALSE}
###### Stacked fungis
# Calc mean over soil&acc for each fungi
sum.fungi.mean.sd <- Topf18.FUNGI %>% group_by(soil, acc) %>%
  summarise(meanApha= mean(Apha, na.rm = TRUE),
            sdApha= sd(Apha, na.rm = TRUE),
            meanFave= mean(Fave, na.rm = TRUE),
            sdFave= sd(Fave, na.rm = TRUE),
            meanFoxy= mean(Foxy, na.rm = TRUE),
            sdFoxy= sd(Foxy, na.rm = TRUE),
            meanFredo= mean(Fredo, na.rm = TRUE),
            sdFredo= sd(Fredo, na.rm = TRUE),
            meanFsol= mean(Fsol, na.rm = TRUE),
            sdFsol= sd(Fsol, na.rm = TRUE),
            meanPhoma= mean(Phoma, na.rm = TRUE),
            sdPhoma= sd(Phoma, na.rm = TRUE),
            meanPult= mean(Pult, na.rm = TRUE),
            sdPult= sd(Pult, na.rm = TRUE),
            meanRsol= mean(Rsol, na.rm = TRUE),
            sdRsol= sd(Rsol, na.rm = TRUE),
            meanAMF= mean(AMF, na.rm = TRUE),
            sdAMF= sd(AMF, na.rm = TRUE),
            meanCros= mean(Cros, na.rm = TRUE),
            sdCros= sd(Cros, na.rm = TRUE))

```



## Analysis of qPCR data  

For the data analysis, I kick-out entries for C2 in F and W, because I only have 1-2 reps per taxa in F soil and only 1 rep for each taxa in W soil.

```{r clean data}
Topf18.FUNGI[Topf18.FUNGI$soil=="F" & Topf18.FUNGI$acc=="C2",7:36] <- NA
Topf18.FUNGI[Topf18.FUNGI$soil=="W" & Topf18.FUNGI$acc=="C2",7:36] <- NA
```

### Plot: Stacks of the fungi

```{r plot stacks, message=FALSE}
# prep data
sum.fungi.accsoil <- Topf18.FUNGI %>% group_by(soil, acc) %>%
  summarise(Apha= mean(Apha, na.rm = TRUE),
            Fave= mean(Fave, na.rm = TRUE),
            Foxy= mean(Foxy, na.rm = TRUE),
            Fredo= mean(Fredo, na.rm = TRUE),
            Fsol= mean(Fsol, na.rm = TRUE),
            Phoma= mean(Phoma, na.rm = TRUE),
            Pult= mean(Pult, na.rm = TRUE),
            Rsol= mean(Rsol, na.rm = TRUE),
            AMF= mean(sqrt(AMF), na.rm = TRUE),    # sqrt AMF!
            Cros= mean(Cros, na.rm = TRUE))

sum.fungi.accsoil$acc <- factor(sum.fungi.accsoil$acc, c("S91", "S134", "G78", "C1", "S64", "G89", "S22", "C2")) # order based on SDWrel in Screen17


sum.fungi.accsoil.long <- reshape2::melt(sum.fungi.accsoil, id = c("soil", "acc")) # put table in long format...
sum.fungi.accsoil.long$variable <- factor(sum.fungi.accsoil.long$variable, levels=c("AMF", "Cros", "Apha", "Fsol", "Foxy", "Rsol", "Pult", "Fave", "Fredo", "Phoma"))
sum.fungi.accsoil.long <- sum.fungi.accsoil.long[order(sum.fungi.accsoil.long$variable), ]  # aus irgend einem grund brauch ich dieses reorder, weil einfach factor() reich nicht...

# Plot
myKolor = c("#A1D99B", "#800026", "darkgreen", "gray22", "gray60", "black", "gray77", "blue", "orange", "#FF3399")

stacks <- ggplot(sum.fungi.accsoil.long, aes(x = acc, fill=variable)) + 
  geom_hline(yintercept = c(-100, 500, 0, 1000, 1500), size = 0.3) +
  geom_bar(data = subset(sum.fungi.accsoil.long, variable %in% c("Fsol", "Apha", "Fave", "Foxy", "Fredo", "Pult", "Rsol", "Phoma")), aes(y = value), stat = "identity") +
  geom_bar(data = subset(sum.fungi.accsoil.long, variable %in% c("AMF", "Cros")), aes(y = -value), stat = "identity") +
  scale_fill_manual(values = myKolor, labels = c("AMF", 
                                                 expression(italic(A.~euteiches)), expression(italic(C.~rosea)),
                                                 expression(italic(F.~avenaceum)), expression(italic(F.~oxysporum)), expression(italic(F.~redolens)),
                                                 expression(italic(F.~solani)), expression(italic(D.~pinodella)), expression(italic(P.~ultimum)),
                                                 expression(italic(R.~solani)) 
                                                 )) +  
  scale_y_continuous(breaks = c(-100,0,500, 1000, 1500)) +
  facet_wrap(~soil, nrow = 1, labeller=labeller(soil = soil_names)) +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.key.size = unit(0.5,"line"),
        legend.text = element_text(size=8),
        legend.text.align = 0,
        strip.background = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=8),
        axis.line = element_blank(),
        axis.text.x = element_text(angle=45, size = 6, vjust=1),
        axis.text.y = element_text(size=5),
        axis.ticks.y = element_blank()) +
  labs(y=expression(paste("Microbial quantities")))

stacks
#ggsave("fungiStack_NS_unfertig.tiff", stacks, units="cm", width=18, height=6, dpi=300, compression = 'lzw')    # I save only once.
rm(stacks)
```

### NMDS

#### First, Data prep: make a DF that contains the mean fungal quants and phenotypic data (SDWRel & RRI) -> "COMBO".


```{r Combo DF}

Topf18.FUNGI.MEANS <- Topf18.FUNGI[,c(1,2,3,4,5,6, 9, 12, 15, 18, 21, 24, 27, 30, 33, 36)]  # this DF contains the mean (between ONE & TWO) qPCR values for each sample only

# then I replace/impute the missing qPCR values
Topf18.FUNGI.MEANS.NONA <- Topf18.FUNGI.MEANS

for (i in thefungis) {
  
  for (sl in levels(Topf18.FUNGI.MEANS.NONA$soil)) {
    for (acs in levels(Topf18.FUNGI.MEANS.NONA$acc)) {
    
       Topf18.FUNGI.MEANS.NONA[Topf18.FUNGI.MEANS.NONA$soil == sl & Topf18.FUNGI.MEANS.NONA$acc == acs,][[i]] <-
         Topf18.FUNGI.MEANS.NONA[Topf18.FUNGI.MEANS.NONA$soil == sl & Topf18.FUNGI.MEANS.NONA$acc == acs,][[i]] %>%
          replace_na(
            mean(Topf18.FUNGI.MEANS.NONA[Topf18.FUNGI.MEANS.NONA$soil == sl & Topf18.FUNGI.MEANS.NONA$acc == acs,][[i]],
                 na.rm = TRUE)
            )
      }
   }
}


# However, C2 in F and W are set to NA for all samples, because, again, I only have 1 (max. 2 in some cases) qPCR observations for theses samples. So I do not want to analyze them.
Topf18.FUNGI.MEANS.NONA[Topf18.FUNGI.MEANS.NONA$soil=="F" & Topf18.FUNGI.MEANS.NONA$acc=="C2",7:16] <- NA
Topf18.FUNGI.MEANS.NONA[Topf18.FUNGI.MEANS.NONA$soil=="W" & Topf18.FUNGI.MEANS.NONA$acc=="C2",7:16] <- NA


# Then prep the actual COMBO DF --> this is the DF I use for the GAM analysis
xxx <- Topf18.FUNGI.MEANS[,-c(2:6)] # throw out extra design variables
Topf18.COMBO.NS <- merge(Topf18.4.8.NS[,c(1:7,17,27)], xxx, "LABEL", all = TRUE)
Topf18.COMBO.NS <- droplevels(Topf18.COMBO.NS)
rm(xxx)

# and do the same for a NONA-COMBO DF  --> this is the DF I use for the NMDS and PERMANOVA analysis
xxx <- Topf18.FUNGI.MEANS.NONA[,-c(2:6)] # throw out extra design variables
Topf18.COMBO.NS.NONA <- merge(Topf18.4.8.NS[,c(1:7,17,27)], xxx, "LABEL", all = TRUE)
Topf18.COMBO.NS.NONA <- droplevels(Topf18.COMBO.NS.NONA)
rm(xxx)

```

...some more prep...
```{r prep data for NMDS}
# Prep data: bring factors in good order for plotting
Topf18.COMBO.NS.NONA$acc <- factor(Topf18.COMBO.NS.NONA$acc, c("C1", "S91", "S134", "S64", "G78", "C2", "S22", "G89"))  # put genotypes in right order for plotting
Topf18.COMBO.NS.NONA$reslevel <- dplyr::recode(Topf18.COMBO.NS.NONA$reslevel, R = "resistant", S = "susceptible")

# prep data for NMDS: I even have to delete the C2-F and C2-W entries entirely for the NMDS analysis (because their all NA)
Topf18.COMBO.NS.NONA.xxx <- subset(Topf18.COMBO.NS.NONA, !(soil=="F" & acc=="C2"))
Topf18.COMBO.NS.NONA.xxx <- subset(Topf18.COMBO.NS.NONA.xxx, !(soil=="W" & acc=="C2"))

```

#### Second, do the NMDS & PERMANOVA: 5 Plots are produced; 1 over 4 soils, as an inset in the 3 soil plot; 1 over 3 soils and then 3 plots for each of the infested soils.... Env.fit arrows are added for the the five fungi with the highest r2 with the ordination. PERMANOVA R2 for the factors soi, acc or reslevel are added as well with corresponding p-vals.

```{r do NMDS, warning=FALSE}
# 1. The 4 soils NMDS Plot without legend -> This willgive a little inset graphic in the big graphic
# prep "Adf", you only want the scores, no factors
Adf4 <- Topf18.COMBO.NS.NONA.xxx[, c(10:19)]
# state the NMDS
fung.mds4 <- metaMDS(Adf4, distance = "bray", k = 2, trymax = 50) # makes the object bci.mds using Bray-Curtis ordination
# pull points from MDS
nmds4.1 <- fung.mds4$points[,1] # also found using scores(birdMDS)
nmds4.2 <- fung.mds4$points[,2]
nmds.plot4 <- cbind(Topf18.COMBO.NS.NONA.xxx, nmds4.1, nmds4.2)
nmds.plot4$acc <- factor(nmds.plot4$acc, c("C1", "S91", "S134", "S64", "G78", "C2", "S22", "G89"))
# plot with ggplot
g4 <- ggplot(nmds.plot4, aes(nmds4.1, nmds4.2)) +
  annotate("text", x=0.5, y=0.6, label= "4 soils", angle = 0, size = 3) +# add fake title inseide Plot
  geom_point(aes(color = soil), size=1.1) +
  scale_color_manual(values = Kolor, labels = soil_nms) +
  scale_fill_manual(values = Kolor) +
  scale_y_continuous(breaks = c(-0.5, 0, 0.5)) +
  scale_x_continuous(breaks = c(-0.5, 0, 0.5)) +
  theme_bw() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = c(0.15,0.2),
        legend.title = element_blank(),
        legend.background = element_rect(fill = "transparent"),
        legend.box.margin = margin(0, 0, 0, 0),
        legend.key.height = unit(0.2,"cm"),
        panel.background =  element_blank(),
        panel.grid = element_blank(),
        plot.title = element_text(size = 10),
        axis.title=element_blank(),
        plot.margin= grid::unit(c(0, 0, 0, 0), "in")) +
  guides(shape=FALSE, fill=FALSE)  #guide_legend(override.aes=list(fill=NA))
  


## 2. 3 infested soils
# prep "Adf", you only want the scores, no factors
Adf3 <- Topf18.COMBO.NS.NONA.xxx[Topf18.COMBO.NS.NONA.xxx$soil!= "F", c(10:19)]
# state the NMDS
fung.mds3 <- metaMDS(Adf3, distance = "bray", k = 2, trymax = 50) # makes the object bci.mds using Bray-Curtis ordination
# pull points from MDS
nmds3.1 <- fung.mds3$points[,1] # also found using scores(birdMDS)
nmds3.2 <- fung.mds3$points[,2]
nmds.plot3 <- cbind(Topf18.COMBO.NS.NONA.xxx[Topf18.COMBO.NS.NONA.xxx$soil!= "F", ], nmds3.1, nmds3.2)

nmds.plot3$acc <- factor(nmds.plot3$acc, c("C1", "S91", "S134", "S64", "G78", "C2", "S22", "G89"))

#PERMANOVA: The r2 and p-vals are then plotted in the plot
fung.dist3 <- vegdist(Adf3)
set.seed(3825)
adonis(fung.dist3 ~ rep + soil*acc, data = Topf18.COMBO.NS.NONA.xxx[Topf18.COMBO.NS.NONA.xxx$soil!= "F", ])
adonis(fung.dist3 ~ rep + soil*reslevel, data = Topf18.COMBO.NS.NONA.xxx[Topf18.COMBO.NS.NONA.xxx$soil!= "F", ])

# Prepare the  10 arrows for the NMDS-Plot
fit3 <- envfit(fung.mds3, Adf3) # "envfit" fittet variabeln auf die Ordination; hier einfach meine 10 Pilze
arrow3 <- data.frame(fit3$vectors$arrows, R = fit3$vectors$r, P = fit3$vectors$pvals)  # Koordinaten der Vektoren der Variabeln
arrow3$FG <- rownames(arrow3)  # Pilznamen

# plot mit ggplot
sze = 2.8  # this is font size 8
g3 <- ggplot(nmds.plot3, aes(nmds3.1, nmds3.2)) +
  geom_segment(data = arrow3[arrow3$R>0.30,], aes(x=0, y=0, xend=NMDS1*arrow3[arrow3$R>0.30,]$R*1.1, yend=NMDS2*arrow3[arrow3$R>0.30,]$R*1.1), # only show the 5 arrows/fungis with the highest R2
               linetype = 1,
               alpha=1,
               size = 0.3,
               arrow = arrow(length = unit(0.15, "cm"))) +
  geom_point(aes(shape = acc, color = soil), size=1.4) +
  scale_shape_manual(values = c(15, 0, 19, 1, 17, 2, 20, 18)) +
  scale_color_manual(values = Kolor[2:4], labels = soil_nms) +
  scale_fill_manual(values = Kolor[2:4]) +
  scale_y_continuous(breaks = c(-0.5, 0, 0.5)) +
  scale_x_continuous(breaks = c(-0.5, 0, 0.5)) +
  coord_cartesian(xlim = c(-0.7,0.7), ylim = c(-0.7,0.7)) +
  annotate("text", x=0.64, y=0.034, label= "AMF", angle = 0, alpha=1, size = 2.8) + # add Fungi names manually
  annotate("text", x=-0.33, y=-0.39, label= "F. sol", fontface = 'italic',  angle = 0,    alpha=1, size = 2.8) +
  annotate("text", x=-0.06, y=-0.43, label= "C. ros.", fontface = 'italic', angle = 0,    alpha=1, size = 2.8) + 
  annotate("text", x=0.155, y=-0.335, label= "F. oxy.", fontface = 'italic',  angle = 0, alpha=1, size = 2.8) + 
  annotate("text", x=-0.39, y=0.15, label= "A. eut.", fontface = 'italic', angle = 0,      alpha=1, size = 2.8) + 
  annotate("text", x=0.45, y=-0.65, label=paste('Stress =', round(fung.mds3$stress,2)), size=2.5) + # add stress-value to plot
  annotate("text", x=-0.7, y=0.7, hjust = 0, label= "PERMANOVA:", size = sze) + 
  annotate("text", x=-0.7, y=0.65, hjust = 0,  label= "~ soil  0.41***", size = sze) +  # PERMANOVA results
  annotate("text", x=-0.7, y=0.6, hjust = 0,  label= "~ soil*acc  0.15***", size = sze) +  
  annotate("text", x=-0.7, y=0.55, hjust = 0, label= "~ acc  0.14***", size = sze) + 
  annotate("text", x=-0.7, y=0.5, hjust = 0, label= "~ reslevel  0.1***", size = sze) +  
  annotate("text", x=-0.7, y=0.45, hjust = 0,  label= "~ soil*reslevel  0.07***", size = sze)+
  theme_bw() +
  theme(legend.position = c(0.15,0.2),
        legend.title = element_blank(),
        legend.background = element_rect(fill = "transparent"),
        legend.box.margin = margin(0, 0, 0, 0),
        legend.key.height = unit(0.2,"cm"),
        panel.background =  element_blank(),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        plot.title = element_text(size = 10)) +
  guides(shape=FALSE, fill=FALSE) +    #guide_legend(override.aes=list(fill=NA))      
  ylab("NMDS2") +
  ggtitle("3 infested soils")


## 3. H soil
# prep "Adf", you only want the scores, no factors
AdfH <- Topf18.COMBO.NS.NONA.xxx[Topf18.COMBO.NS.NONA.xxx$soil== "H", c(10,12,14:19)]   # take out Fave & Fredo in H 
# state the NMDS
fung.mdsH <- metaMDS(AdfH, distance = "bray", k = 2, trymax = 50) #makes the object bci.mds using Bray-Curtis ordination
# pull points from MDS
nmdsH.1 <- fung.mdsH$points[,1] # also found using scores(birdMDS)
nmdsH.2 <- fung.mdsH$points[,2]
nmds.plotH <- cbind(Topf18.COMBO.NS.NONA.xxx[Topf18.COMBO.NS.NONA.xxx$soil== "H", c(1:10,12,14:19)], nmdsH.1, nmdsH.2)   # take out Fave & Fredo in H 

nmds.plotH$acc <- factor(nmds.plotH$acc, c("C1", "S91", "S134", "S64", "G78", "C2", "S22", "G89"))

#PERMANOVA: The r2 and p-vals are then plotted in the plot
fung.distH <- vegdist(AdfH)
set.seed(3825)
adonis(fung.distH ~ rep + acc, data = Topf18.COMBO.NS.NONA.xxx[Topf18.COMBO.NS.NONA.xxx$soil== "H", ])
adonis(fung.distH ~ rep + reslevel, data = Topf18.COMBO.NS.NONA.xxx[Topf18.COMBO.NS.NONA.xxx$soil== "H", ])

# Prepare the  10 arrows for the NMDS-Plot
fitH <- envfit(fung.mdsH, AdfH) # "envfit" fittet variabeln auf die Ordination; hier einfach meine 10 Pilze
arrowH <- data.frame(fitH$vectors$arrows, R = fitH$vectors$r, P = fitH$vectors$pvals)  # Koordinaten der Vektoren der Variabeln
arrowH$FG <- rownames(arrowH)  # Pilznamen

# plot mit ggplot
gH <- ggplot(nmds.plotH, aes(nmdsH.1, nmdsH.2)) +
  geom_segment(data = arrowH[arrowH$R>0.22,], aes(x=0, y=0, xend=NMDS1*arrowH[arrowH$R>0.22,]$R*1.1, yend=NMDS2*arrowH[arrowH$R>0.22,]$R*1.1), # only show the 5 arrows/fungis with the highest R2
               linetype = 1,
               alpha=1,
               size = 0.3,
               arrow = arrow(length = unit(0.15, "cm"))) +
  geom_point(aes(shape = acc), color = Kolor[2], fill= Kolor[2], size=1.4) +
  scale_shape_manual(values = c(17, 19, 15, 20, 18, 0, 1, 2)) +
  stat_ellipse(aes(lty=factor(reslevel)), alpha=0.2, size = 0.5, geom = "polygon", color = Kolor[2], fill=NA) +
  scale_y_continuous(breaks = c(-0.5, 0, 0.5)) +
  scale_x_continuous(breaks = c(-0.5, 0, 0.5)) +
  coord_cartesian(xlim = c(-0.7,0.7), ylim = c(-0.7,0.7)) +
  annotate("text", x=-0.08, y=0.74, label= "A. eut", fontface = 'italic', angle = 0, alpha=1, size = 2.8) + 
  annotate("text", x=0.41, y=0.055, label= "F. oxy.", fontface = 'italic', angle = 0,    alpha=1, size = 2.8) + 
  annotate("text", x=0.6, y=0.02, label= "F. sol.", fontface = 'italic',  angle = 0, alpha=1, size = 2.8) + 
  annotate("text", x=-0.56, y=-0.41, label= "AMF",                          angle = 0,    alpha=1, size = 2.8) +
  annotate("text", x=0.47, y=-0.088, label= "C. ros.", fontface = 'italic', angle = 0,      alpha=1, size = 2.8) + 
  annotate("text", x=-0.7, y=0.7, hjust = 0, label= "PERMANOVA:", size = sze) + 
  annotate("text", x=-0.7, y=0.65, hjust = 0, label= "~ acc  0.59***", size = sze) + 
  annotate("text", x=-0.7, y=0.6, hjust = 0, label= "~ reslevel  0.39***", size = sze) +  
  annotate("text", x=0.45, y=-0.65, label=paste('Stress =', round(fung.mdsH$stress,2)), size=2.5) +
  ggtitle("Kirchlindach") +
  theme_bw() +
  theme(legend.title = element_blank(),
        legend.background = element_rect(fill = "transparent"),
        legend.box.margin = margin(0, 0, 0, 0),
        legend.key.height = unit(0.2,"cm"),
        panel.background =  element_blank(),
        panel.grid = element_blank(),
        plot.title = element_text(size = 10),
        axis.title = element_blank(),
        legend.spacing.x = unit(0.02, 'cm'),
        legend.spacing.y = unit(0.02, 'cm')) +
  guides(fill = guide_legend(colour="black"),
         shape = guide_legend(override.aes = list(shape = c(17, 19, 15, 20, 18, 0, 1, 2), color="black", fill="black")),
         lty = guide_legend(override.aes = list(color="black")))



## 4. L soil
# prep "Adf", you only want the scores, no factors
AdfL <- Topf18.COMBO.NS.NONA.xxx[Topf18.COMBO.NS.NONA.xxx$soil== "L", c(10:12,14:19)]   # take out Fredo in L 
# state the NMDS
fung.mdsL <- metaMDS(AdfL, distance = "bray", k = 2, trymax = 50) #makes the object bci.mds using Bray-Curtis ordination
# pull points from MDS
nmdsL.1 <- fung.mdsL$points[,1] # also found using scores(birdMDS)
nmdsL.2 <- fung.mdsL$points[,2]
nmds.plotL <- cbind(Topf18.COMBO.NS.NONA.xxx[Topf18.COMBO.NS.NONA.xxx$soil== "L", c(1:12, 14:19)], nmdsL.1, nmdsL.2)  # take out Fredo in L 

nmds.plotL$acc <- factor(nmds.plotL$acc, c("C1", "S91", "S134", "S64", "G78", "C2", "S22", "G89"))

#PERMANOVA: The r2 and p-vals are then plotted in the plot
fung.distL <- vegdist(AdfL)
set.seed(3825)
adonis(fung.distL ~ rep + acc, data = Topf18.COMBO.NS.NONA.xxx[Topf18.COMBO.NS.NONA.xxx$soil== "L", ])
adonis(fung.distL ~ rep + reslevel, data = Topf18.COMBO.NS.NONA.xxx[Topf18.COMBO.NS.NONA.xxx$soil== "L", ])

# Prepare the  10 arrows for the NMDS-Plot
fitL <- envfit(fung.mdsL, AdfL) # "envfit" fittet variabeln auf die Ordination; hier einfach meine 10 Pilze
arrowL <- data.frame(fitL$vectors$arrows, R = fitL$vectors$r, P = fitL$vectors$pvals)  # Koordinaten der Vektoren der Variabeln
arrowL$FG <- rownames(arrowL)  # Pilznamen

# plot mit ggplot
gL <- ggplot(nmds.plotL, aes(nmdsL.1, nmdsL.2)) +
  geom_segment(data = arrowL[arrowL$R>0.47,], aes(x=0, y=0, xend=NMDS1*arrowL[arrowL$R>0.47,]$R*1.1, yend=NMDS2*arrowL[arrowL$R>0.47,]$R*1.1), # only show the 5 arrows/fungis with the highest R2
               linetype = 1,
               alpha=1,
               size = 0.3,
               arrow = arrow(length = unit(0.15, "cm"))) +
  geom_point(aes(shape = acc), color = Kolor[3], fill= Kolor[3], size=1.4) +
  scale_shape_manual(values = c(17, 19, 15, 20, 18, 0, 1, 2)) +
  stat_ellipse(aes(lty=factor(reslevel)), alpha=0.2, size = 0.5, geom = "polygon", color = Kolor[3], fill=NA) +
  scale_y_continuous(breaks = c(-0.5, 0, 0.5)) +
  scale_x_continuous(breaks = c(-0.5, 0, 0.5)) +
  coord_cartesian(xlim = c(-0.7,0.7), ylim = c(-0.7,0.7)) +
  annotate("text", x=-0.32, y=-0.43, label= "A. eut.", fontface = 'italic',  alpha=1, size = 2.8) + 
  annotate("text", x=-0.25, y=-0.55, label= "F. oxy.", fontface = 'italic', alpha=1, size = 2.8) +
  annotate("text", x=-0.65, y=-0.04, label= "F sol.", fontface = 'italic', alpha=1, size = 2.8) + 
  annotate("text", x=0.1, y=-0.6, label= "P. ult.", fontface = 'italic', alpha=1, size = 2.8) + 
  annotate("text", x=0.68, y=-0.29, label= "AMF",   alpha=1, size = 2.8) + # add Fungi names manually
  annotate("text", x=-0.7, y=0.7, hjust = 0, label= "PERMANOVA:", size = sze) + 
  annotate("text", x=-0.7, y=0.65, hjust = 0, label= "~ acc  0.46**", size = sze) + 
  annotate("text", x=-0.7, y=0.6, hjust = 0, label= "~ reslevel  0.21**", size = sze) +  
  annotate("text", x=0.45, y=-0.65, label=paste('Stress =', round(fung.mdsL$stress,2)), size=2.5) +
  ggtitle("Puch") +
  theme_bw() +
  theme(legend.title = element_blank(),
        legend.background = element_rect(fill = "transparent"),
        legend.box.margin = margin(0, 0, 0, 0),
        legend.key.height = unit(0.2,"cm"),
        panel.background =  element_blank(),
        panel.grid = element_blank(),
        plot.title = element_text(size = 10),
        legend.spacing.x = unit(0.02, 'cm'),
        legend.spacing.y = unit(0.02, 'cm')) +
  guides(fill = guide_legend(colour="black"),
         shape = guide_legend(override.aes = list(shape = c(17, 19, 15, 20, 18, 0, 1, 2), color="black", fill="black")),
         lty = guide_legend(override.aes = list(color="black"))) +
  xlab("NMDS1") +
  ylab("NMDS2")



## 5. W soil
# prep "Adf", you only want the scores, no factors
AdfW <- Topf18.COMBO.NS.NONA.xxx[Topf18.COMBO.NS.NONA.xxx$soil== "W", c(10, 12:19)]   # take out Fave in W 
# state the NMDS
fung.mdsW <- metaMDS(AdfW, distance = "bray", k = 2, trymax = 50) #makes the object bci.mds using Bray-Curtis ordination
# pull points from MDS
nmdsW.1 <- fung.mdsW$points[,1] # also found using scores(birdMDS)
nmdsW.2 <- fung.mdsW$points[,2]
nmds.plotW <- cbind(Topf18.COMBO.NS.NONA.xxx[Topf18.COMBO.NS.NONA.xxx$soil== "W", c(1:10, 12:19)], nmdsW.1, nmdsW.2)  # take out Fave in W 

nmds.plotW$acc <- factor(nmds.plotW$acc, c("C1", "S91", "S134", "S64", "G78", "C2", "S22", "G89"))

#PERMANOVA: The r2 and p-vals are then plotted in the plot
fung.distW <- vegdist(AdfW)
set.seed(3825)
adonis(fung.distW ~ rep + acc, data = Topf18.COMBO.NS.NONA.xxx[Topf18.COMBO.NS.NONA.xxx$soil== "W", ])
adonis(fung.distW ~ rep + reslevel, data = Topf18.COMBO.NS.NONA.xxx[Topf18.COMBO.NS.NONA.xxx$soil== "W", ])

# Prepare the  10 arrows for the NMDS-Plot
fitW <- envfit(fung.mdsW, AdfW) # "envfit" fittet variabeln auf die Ordination; hier einfach meine 10 Pilze
arrowW <- data.frame(fitW$vectors$arrows, R = fitW$vectors$r, P = fitW$vectors$pvals)  # Koordinaten der Vektoren der Variabeln
arrowW$FG <- rownames(arrowW)  # Pilznamen
  
# plot mit ggplot
gW <- ggplot(nmds.plotW, aes(nmdsW.1, nmdsW.2)) +
 geom_segment(data = arrowW[arrowW$R>0.32,], aes(x=0, y=0, xend=NMDS1*arrowW[arrowW$R>0.32,]$R*1.1, yend=NMDS2*arrowW[arrowW$R>0.32,]$R*1.1), # only show the 5 arrows/fungis with the highest R2
              linetype = 1,
              alpha=1,
              size = 0.3,
              arrow = arrow(length = unit(0.15, "cm"))) +
 geom_point(aes(shape = acc), color = Kolor[4], fill= Kolor[4], size=1.4) +
 scale_shape_manual(values = c(17, 19, 15, 20, 18, 1, 2)) +
 stat_ellipse(aes(lty=factor(reslevel)), alpha=0.2, size = 0.5, geom = "polygon", color = Kolor[4], fill=NA) +
 scale_y_continuous(breaks = c(-0.5, 0, 0.5)) +
 scale_x_continuous(breaks = c(-0.5, 0, 0.5)) +
 coord_cartesian(xlim = c(-0.7,0.7), ylim = c(-0.7,0.7)) +
 annotate("text", x=-0.25, y=-0.6, label= "A. eut.", fontface = 'italic', alpha=1, size = 2.8) + # add Fungi names manually
 annotate("text", x=0.4, y=-0.1, label= "F. oxy.", fontface = 'italic', alpha=1, size = 2.8) + 
 annotate("text", x=0.38, y=-0.4, label= "F. redo.", fontface = 'italic', alpha=1, size = 2.8) +
 annotate("text", x=0.16, y=-0.39, label= "F. sol.", fontface = 'italic', alpha=1, size = 2.8) + 
 annotate("text", x=0.33, y=-0.25, label= "C. ros.", fontface = 'italic', alpha=1, size = 2.8) + 
 annotate("text", x=-0.7, y=0.7, hjust = 0, label= "PERMANOVA:", size = sze) + 
 annotate("text", x=-0.7, y=0.65, hjust = 0, label= "~ acc  0.33*", size = sze) + 
 annotate("text", x=-0.7, y=0.6, hjust = 0, label= "~ reslevel  0.18**", size = sze) +  
 annotate("text", x=0.45, y=-0.65, label=paste('Stress =', round(fung.mdsW$stress,2)), size=2.5) +
 ggtitle("Neu-Eichenberg") +
 theme_bw() +
 theme(legend.title = element_blank(),
       legend.background = element_rect(fill = "transparent"),
       legend.box.margin = margin(0, 0, 0, 0),
       legend.key.height = unit(0.2,"cm"),
       panel.background =  element_blank(),
       panel.grid = element_blank(),
       plot.title = element_text(size = 10),
       axis.title.y = element_blank(),
       legend.spacing.x = unit(0.02, 'cm'),
       legend.spacing.y = unit(0.02, 'cm')) +
 guides(fill = guide_legend(colour="black"),
        shape = guide_legend(override.aes = list(shape = c(17, 19, 15, 20, 18, 1, 2), color="black", fill="black")),
        lty = guide_legend(override.aes = list(color="black"))) +
  xlab("NMDS1")


## 6. Now, put all plot together
# A: get legends (note the 4 soils legend is already in the big panel), and arranghe them together
thelegendsoils <- ggpubr::get_legend(g4 + theme(legend.margin = margin(0, 0, 0, 0)) + guides(lty=FALSE)) # extract the legend for the soils from the 4 soils plot
thelegendAccs <- ggpubr::get_legend(gH + theme(legend.margin = margin(0, 0, 0, 0)) + guides(lty=FALSE)) # extract the legend for the accs from one of the plots
thelegendRes <- ggpubr::get_legend(gH + theme(legend.margin = margin(0, 0, 0, 0)) +
                                     guides(shape=FALSE, lty = guide_legend(override.aes = list(color ="black"))))
legendS <- ggpubr::ggarrange(thelegendsoils, thelegendAccs, thelegendRes) # 

# B: produce the 4 soils inlet in the 3 soils plot
g4_grob <- ggplotGrob(g4 + theme(legend.position = "none"))
g3plus <- g3 + annotation_custom(grob = g4_grob, xmin = 0.3, xmax = 0.75, ymin = 0.3, ymax = 0.75)

# C: arrange the 4 plots
wow <- cowplot::plot_grid(
  g3plus + theme(legend.position="none") + theme(axis.text.x = element_blank(), axis.title = element_text(size=8)),
  gH + theme(legend.position="none") + theme(axis.text = element_blank()),
  gL + theme(legend.position="none") + theme(axis.title = element_text(size=8)),
  gW + theme(legend.position="none") + theme(axis.text.y = element_blank(), axis.title = element_text(size=8)),
  nrow = 2,
  rel_widths = c(1.11, 1, 1.11, 1))

wow

# D: save that beautiful plot & the legend!
#ggsave("4plus.tiff", wow, units="cm", width=18, height=18, dpi=300, compression = 'lzw')   # 159mm breit ist standard Word doc. 190mm breit ist Frontiers in Plant Sci.
#ggsave("4plusLEGEND.tiff", legendS, units="cm", width=6, height=6, dpi=300, compression = 'lzw')


#rm(Adf4, Adf3, AdfH, AdfL, AdfW, arrow3, arrowH, arrowL, arrowW, fit3, fitH, fitL, fitW, fung.mds3, fung.mdsH, fung.mds4, fung.mdsL, fung.mdsW, fung.dist3, fung.distH, fung.distL, fung.distW, nmds.plot4, nmds.plot3, nmds.plotH, nmds.plotL, nmds.plotW, fung.dist3, fung.distH,fung.distL, fung.distW, nmds4.1, nmds4.2, nmds3.1, nmds3.2, nmdsH.1, nmdsH.2, nmdsL.1, nmdsL.2, nmdsW.1, nmdsW.2, g4, g4_grob, g3, g3plus, gH, gL, gW)
```


### GAM Analysis of SDWrel. ~ fungi

1) I backward select the variables (fungi) from a full model containing all fungi, without factors soil or acc. At each step I exlude the variable with the highest p-value. I set k=5 manually, in order to avoid too much wiggeliness. 


```{r GAM 100noacc}

d3ta <- Topf18.COMBO.NS %>% filter (soil!="F")

modGAM_100 <- gam(s.dw.ratioSNS ~ 
                                  s(Apha, bs="cs", k=5) +
                                  s(Fsol, bs="cs", k=5) +
                                  s(Fave, bs="cs", k=5) +
                                  s(Fredo, bs="cs", k=5) +
                                  s(Foxy, bs="cs", k=5) +
                                  s(Phoma, bs="cs", k=5) +
                                  s(Pult, bs="cs", k=5) +
                                  s(Rsol, bs="cs", k=5) +
                                  s(Cros, bs="cs", k=5) +
                                  s(AMF, bs="cs", k=5),
                                      data = d3ta)  

#
modelxxx <- modGAM_100
# model diagnostic
par(mfrow = c(2,2))
gam.check(modelxxx)
# model inspection
plot.gam(modelxxx, pages=1, scheme = 2)
summary.gam(modelxxx)
AIC(modelxxx)

```
-> Pult rausnehmen:

```{r GAM 99noacc1}
modGAM_99 <- gam(s.dw.ratioSNS ~ 
                                  s(Apha, bs="cs", k=5) +
                                  s(Fsol, bs="cs", k=5) +
                                  s(Fave, bs="cs", k=5) +
                                  s(Fredo, bs="cs", k=5) +
                                  s(Foxy, bs="cs", k=5) +
                                  s(Phoma, bs="cs", k=5) +
                                  #s(Pult, bs="cs", k=5) +
                                  s(Rsol, bs="cs", k=5) +
                                  s(Cros, bs="cs", k=5) +
                                  s(AMF, bs="cs", k=5),
                                      data = d3ta)  

#
modelxxx <- modGAM_99
# model diagnostic
par(mfrow = c(2,2))
gam.check(modelxxx)
# model inspection
plot.gam(modelxxx, pages=1, scheme = 2)
summary.gam(modelxxx)
AIC(modelxxx)
```

-> Phoma rausnehmen:

```{r GAM 99noacc}
modGAM_98 <- gam(s.dw.ratioSNS ~ 
                                  s(Apha, bs="cs", k=5) +
                                  s(Fsol, bs="cs", k=5) +
                                  s(Fave, bs="cs", k=5) +
                                  s(Fredo, bs="cs", k=5) +
                                  s(Foxy, bs="cs", k=5) +
                                  #s(Phoma, bs="cs", k=5) +
                                  #s(Pult, bs="cs", k=5) +
                                  s(Rsol, bs="cs", k=5) +
                                  s(Cros, bs="cs", k=5) +
                                  s(AMF, bs="cs", k=5),
                                      data = d3ta)  

#
modelxxx <- modGAM_98
# model diagnostic
par(mfrow = c(2,2))
gam.check(modelxxx)
# model inspection
plot.gam(modelxxx, pages=1, scheme = 2)
summary.gam(modelxxx)
AIC(modelxxx)
```

-> Phoma rausnehmen:

```{r GAM 98noacc}
modGAM_98 <- gam(s.dw.ratioSNS ~ 
                                  s(Apha, bs="cs", k=5) +
                                  s(Fsol, bs="cs", k=5) +
                                  s(Fave, bs="cs", k=5) +
                                  s(Fredo, bs="cs", k=5) +
                                  s(Foxy, bs="cs", k=5) +
                                  #s(Phoma, bs="cs", k=5) +
                                  #s(Pult, bs="cs", k=5) +
                                  s(Rsol, bs="cs", k=5) +
                                  s(Cros, bs="cs", k=5) +
                                  s(AMF, bs="cs", k=5),
                                      data = d3ta)  

#
modelxxx <- modGAM_98
# model diagnostic
par(mfrow = c(2,2))
gam.check(modelxxx)
# model inspection
plot.gam(modelxxx, pages=1, scheme = 2)
summary.gam(modelxxx)
AIC(modelxxx)
```

-> take Foxy out: 


```{r GAM 97noacc}
modGAM_97 <- gam(s.dw.ratioSNS ~ 
                                  s(Apha, bs="cs", k=5) +
                                  s(Fsol, bs="cs", k=5) +
                                  s(Fave, bs="cs", k=5) +
                                  s(Fredo, bs="cs", k=5) +
                                  #s(Foxy, bs="cs", k=5) +
                                  #s(Phoma, bs="cs", k=5) +
                                  #s(Pult, bs="cs", k=5) +
                                  s(Rsol, bs="cs", k=5) +
                                  s(Cros, bs="cs", k=5) +
                                  s(AMF, bs="cs", k=5),
                                      data = d3ta)  

#
modelxxx <- modGAM_97
# model diagnostic
par(mfrow = c(2,2))
gam.check(modelxxx)
# model inspection
plot.gam(modelxxx, pages=1, scheme = 2)
summary.gam(modelxxx)
AIC(modelxxx)
```

-> take Fave out:

```{r GAM 96noacc}
modGAM_96 <- gam(s.dw.ratioSNS ~ 
                                  s(Apha, bs="cs", k=5) +
                                  s(Fsol, bs="cs", k=5) +
                                  #s(Fave, bs="cs", k=5) +
                                  s(Fredo, bs="cs", k=5) +
                                  #s(Foxy, bs="cs", k=5) +
                                  #s(Phoma, bs="cs", k=5) +
                                  #s(Pult, bs="cs", k=5) +
                                  s(Rsol, bs="cs", k=5) +
                                  s(Cros, bs="cs", k=5) +
                                  s(AMF, bs="cs", k=5),
                                      data = d3ta)  

#
modelxxx <- modGAM_96
# model diagnostic
par(mfrow = c(2,2))
gam.check(modelxxx)
# model inspection
plot.gam(modelxxx, pages=1, scheme = 2)
summary.gam(modelxxx)
AIC(modelxxx)
```
-> finally, Apha is eliminated from the model:

```{r GAM 95noacc}
modGAM_95 <- gam(s.dw.ratioSNS ~ 
                                  #s(Apha, bs="cs", k=5) +
                                  s(Fsol, bs="cs", k=5) +
                                  #s(Fave, bs="cs", k=5) +
                                  s(Fredo, bs="cs", k=5) +
                                  #s(Foxy, bs="cs", k=5) +
                                  #s(Phoma, bs="cs", k=5) +
                                  #s(Pult, bs="cs", k=5) +
                                  s(Rsol, bs="cs", k=5) +
                                  s(Cros, bs="cs", k=5) +
                                  s(AMF, bs="cs", k=5),
                                      data = d3ta)  

#
modelxxx <- modGAM_95
# model diagnostic
par(mfrow = c(2,2))
gam.check(modelxxx)
# model inspection
plot.gam(modelxxx, pages=1, scheme = 2)
summary.gam(modelxxx)
AIC(modelxxx)
concurvity(modelxxx)
```
So, in the final model over three infested soils I have the three pathogens Fsol, Fredo & Rsol that have sig. smooth terms and the two beneficials AMF & Cros. This model has an R2=0.54, dev. expl. 58.6%, GCV=0.026. There seems to be a bit of concurvity problem.  

Now, claculate adj.R2 for each of the retained variables:  
-> I will use these values to display them in the plots further down.


```{r GAM mod 95noacc factors contributions}

R2_Fsol <- unlist(summary(modGAM_95)[c(10,14)]) - 
          unlist(summary(gam(s.dw.ratioSNS ~  #s(Fsol, bs="cs", k=5) +
                                s(Fredo, bs="cs", k=5) + s(Rsol, bs="cs", k=5) + s(Cros, bs="cs", k=5) + s(AMF, bs="cs", k=5), data = d3ta))[c(10,14)])

R2_Fredo <- unlist(summary(modGAM_95)[c(10,14)]) - 
          unlist(summary(gam(s.dw.ratioSNS ~  s(Fsol, bs="cs", k=5) + #s(Fredo, bs="cs", k=5) +
                                s(Rsol, bs="cs", k=5) + s(Cros, bs="cs", k=5) + s(AMF, bs="cs", k=5), data = d3ta))[c(10,14)])

R2_Rsol <- unlist(summary(modGAM_95)[c(10,14)]) - 
          unlist(summary(gam(s.dw.ratioSNS ~  s(Fsol, bs="cs", k=5) + s(Fredo, bs="cs", k=5) + #s(Rsol, bs="cs", k=5) + 
                               s(Cros, bs="cs", k=5) + s(AMF, bs="cs", k=5), data = d3ta))[c(10,14)])

R2_Cros <- unlist(summary(modGAM_95)[c(10,14)]) - 
          unlist(summary(gam(s.dw.ratioSNS ~  s(Fsol, bs="cs", k=5) + s(Fredo, bs="cs", k=5) + s(Rsol, bs="cs", k=5) + #s(Cros, bs="cs", k=5) + 
                               s(AMF, bs="cs", k=5), data = d3ta))[c(10,14)])

R2_AMF <-  unlist(summary(modGAM_95)[c(10,14)]) - 
          unlist(summary(gam(s.dw.ratioSNS ~  s(Fsol, bs="cs", k=5) + s(Fredo, bs="cs", k=5) + s(Rsol, bs="cs", k=5) + s(Cros, bs="cs", k=5) #s(AMF, bs="cs", k=5)
                               , data = d3ta))[c(10,14)])

(R2s <- as.data.frame(rbind(R2_Fsol, R2_Fredo, R2_Rsol, R2_Cros, R2_AMF)))
sum(R2s$dev.expl)
sum(R2s$r.sq)

```

#### Now, we redo the same procedure for each of the three infested soils individually:

```{r GAM soil H 100}
modGAM_H_100 <- gam(s.dw.ratioSNS ~ 
                                  3s(Apha, bs="cs", k=5) +
                                  s(Fsol, bs="cs", k=5)  +
                                  #s(Fave, bs="cs", k=5) +
                                  #s(Fredo, bs="cs", k=5) +
                                  #s(Foxy, bs="cs", k=5) +   # Foxy is not in the model right from the start cause it shows high concurvity with Fsol
                                  #s(Phoma, bs="cs", k=5) +
                                  #s(Pult, bs="cs", k=5) +
                                  #s(Rsol, bs="cs", k=5) +
                                  s(Cros, bs="cs", k=5),
                                  #s(AMF, bs="cs", k=5),
                                      data = d3ta[d3ta$soil=="H",])  

#
modelxxx <- modGAM_H_100
# model diagnostic
par(mfrow = c(2,2))
gam.check(modelxxx)
# model inspection
plot.gam(modelxxx, pages=1, scheme = 2)
summary.gam(modelxxx)
AIC(modelxxx)
concurvity(modelxxx)

# contribution - r2
R2_Fsol <- unlist(summary(modGAM_H_100)[c(10,14)]) - 
          unlist(summary(gam(s.dw.ratioSNS ~  #s(Fsol, bs="cs", k=5) +
                               s(Cros, bs="cs", k=5), data = d3ta[d3ta$soil=="H",]))[c(10,14)])

R2_Cros <- unlist(summary(modGAM_H_100)[c(10,14)]) - 
          unlist(summary(gam(s.dw.ratioSNS ~  s(Fsol, bs="cs", k=5), #s(Cros, bs="cs", k=5),
                             data = d3ta[d3ta$soil=="H",]))[c(10,14)])

(R2s <- as.data.frame(rbind(R2_Fsol, R2_Cros)))
sum(R2s$dev.expl)
sum(R2s$r.sq)
```
In H only Fsol and Cros are retained.

```{r GAM soil L 100}
modGAM_L_100 <- gam(s.dw.ratioSNS ~ 
                                  #s(Apha, bs="cs", k=5) +
                                  s(Fsol, bs="cs", k=5)  +
                                  #s(Fave, bs="cs", k=5) +
                                  #s(Fredo, bs="cs", k=5l) +
                                  #s(Foxy, bs="cs", k=5) +
                                  #s(Phoma, bs="cs", k=5) +
                                  #s(Pult, bs="cs", k=5) +
                                  #s(Rsol, bs="cs", k=5) +
                                  #s(Cros, bs="cs", k=5) +
                                  s(AMF, bs="cs", k=5),
                                      data = d3ta[d3ta$soil=="L",])
#
modelxxx <- modGAM_L_100
# model diagnostic
par(mfrow = c(2,2))
gam.check(modelxxx)
# model inspection
plot.gam(modelxxx, pages=1, scheme = 2)
summary.gam(modelxxx)
AIC(modelxxx)
concurvity(modelxxx)

# contribution - r2


R2_Fsol <- unlist(summary(modGAM_L_100)[c(10,14)]) - 
          unlist(summary(gam(s.dw.ratioSNS ~  #s(Fsol, bs="cs", k=5) +
                               s(AMF, bs="cs", k=5), data = d3ta[d3ta$soil=="L",]))[c(10,14)])

R2_AMF <- unlist(summary(modGAM_L_100)[c(10,14)]) - 
          unlist(summary(gam(s.dw.ratioSNS ~  s(Fsol, bs="cs", k=5) #+ s(AMF, bs="cs", k=5)
                             , data = d3ta[d3ta$soil=="L",]))[c(10,14)])


(R2s <- as.data.frame(rbind(R2_Fsol, R2_AMF)))
sum(R2s$dev.expl)
sum(R2s$r.sq)

```
In L soil Fsol & AMF are retained.  
Foxy has a sign. smooth term but produces considerable concurvity when in the Model.

```{r GAM soil W 100}
modGAM_W_100 <- gam(s.dw.ratioSNS ~ 
                                  #s(Apha, bs="cs", k=5) +
                                  s(Fsol, bs="cs", k=5)  +
                                  #s(Fave, bs="cs", k=5) +
                                  #s(Fredo, bs="cs", k=5) +
                                  #s(Foxy, bs="cs", k=5) +
                                  #s(Phoma, bs="cs", k=5) +
                                  #s(Pult, bs="cs", k=5) +
                                  s(Rsol, bs="cs", k=5) +
                                  #s(Cros, bs="cs", k=5) +
                                  s(AMF, bs="cs", k=5),
                                      data = d3ta[d3ta$soil=="W",])  

#
modelxxx <- modGAM_W_100
# model diagnostic
par(mfrow = c(2,2))
gam.check(modelxxx)
# model inspection
plot.gam(modelxxx, pages=1, scheme = 2)
summary.gam(modelxxx)
AIC(modelxxx)

# contribution - r2
R2_Fsol <- unlist(summary(modGAM_W_100)[c(10,14)]) - 
          unlist(summary(gam(s.dw.ratioSNS ~  #s(Fsol, bs="cs", k=5) +
                               s(Rsol, bs="cs", k=5) + s(AMF, bs="cs", k=5), data = d3ta[d3ta$soil=="W",]))[c(10,14)])

R2_Rsol <- unlist(summary(modGAM_W_100)[c(10,14)]) - 
          unlist(summary(gam(s.dw.ratioSNS ~  s(Fsol, bs="cs", k=5) + #s(Rsol, bs="cs", k=5) +
                               s(AMF, bs="cs", k=5), data = d3ta[d3ta$soil=="W",]))[c(10,14)])

R2_AMF <- unlist(summary(modGAM_W_100)[c(10,14)]) - 
          unlist(summary(gam(s.dw.ratioSNS ~  s(Fsol, bs="cs", k=5) + s(Rsol, bs="cs", k=5) #+ s(AMF, bs="cs", k=5)
                             , data = d3ta[d3ta$soil=="W",]))[c(10,14)])


(R2s <- as.data.frame(rbind(R2_Fsol, R2_Rsol, R2_AMF)))
sum(R2s$dev.expl)
sum(R2s$r.sq)
```

In W Fsol, Rsol & AMF are retained.


#### Plot the GAM plots on the original scale of the respons variable

```{r  gam plot on original scale, warning=FALSE}
#setup
si <- 4
sz <- 3
axsz <- 8
axtk <- 6

# 3 soils
gp_3_Fsol <- tidymv::plot_smooths(model = modGAM_95, series = Fsol) +
  theme_bw() +
  geom_point(data=d3ta, aes(x=Fsol, y=s.dw.ratioSNS, color=soil), size=2) +
  annotate("text", x=1600, y=1.25, label= expression(italic("F. solani")), size = si) +
  annotate("text", x=1600, y=1.1, label= expression(paste(italic(r^2), "= 0.07")) , size = sz) +
  scale_color_manual(values=Kolor[2:4]) +
  scale_y_continuous(breaks = c(0.5, 1)) +
    labs(y=expression(SDW[italic(Rel.)]), x="DNA amount") +
  theme(legend.position = "none",
        axis.title =element_text(size=axsz),
        axis.text = element_text(size=axtk),
          panel.grid = element_blank()) +
  coord_cartesian(ylim = c(0, 1.3))
  
gp_3_Fredo <- tidymv::plot_smooths(model = modGAM_95, series = Fredo) +
  theme_bw() +
  geom_point(data=d3ta, aes(x=Fredo, y=s.dw.ratioSNS, color=soil), size=2) +
  annotate("text", x=15, y=1.25, label= expression(italic("F. redolens")), size = si) +
  annotate("text", x=1500, y=1.1, label= expression(paste(italic(r^2), "= 0.05")), size = sz) +
  scale_color_manual(values=Kolor[2:4])+
  scale_y_continuous(breaks = c(0.5, 1)) +
  labs(y=expression(SDW[italic(Rel.)])) +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.title =element_text(size=axsz),
        axis.text = element_text(size=axtk),
          panel.grid = element_blank()) +
  coord_cartesian(ylim = c(0, 1.3)) 

gp_3_Rsol <- tidymv::plot_smooths(model = modGAM_95, series = Rsol) +
  theme_bw() +
  geom_point(data=d3ta, aes(x=Rsol, y=s.dw.ratioSNS, color=soil),  size=2) +
  annotate("text", x=150, y=1.25, label= expression(italic("R. solani")),  size = si) +
  annotate("text", x=150, y=1.1, label= expression(paste(italic(r^2), "= 0.04")), size = sz) +
  scale_color_manual(values=Kolor[2:4])+
  scale_y_continuous(breaks = c(0.5, 1)) +
  labs(y=expression(SDW[italic(Rel.)])) +
  theme(legend.position = "none",
        axis.title =element_text(size=axsz),
        axis.title.x = element_blank(),
        axis.text = element_text(size=axtk),
          panel.grid = element_blank()) +
  coord_cartesian(ylim = c(0, 1.3)) 

gp_3_Cros <- tidymv::plot_smooths(model = modGAM_95, series = Cros) +
  theme_bw() +
  geom_point(data=d3ta, aes(x=Cros, y=s.dw.ratioSNS, color=soil),  size=2) +
  labs(y=expression(SDW[italic(Rel.)])) +
  annotate("text", x=60, y=1.25, label= expression(italic("C. rosea")), size = si) +
  annotate("text", x=60, y=1.1, label= expression(paste(italic(r^2), "= 0.04")), size = sz) + 
  scale_color_manual(values=Kolor[2:4])+
  scale_y_continuous(breaks = c(0.5, 1)) +
    labs(y=expression(SDW[italic(Rel.)])) +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.title =element_text(size=axsz),
        axis.text = element_text(size=axtk),
          panel.grid = element_blank()) +
  coord_cartesian(ylim = c(0, 1.3)) 

gp_3_AMF <- tidymv::plot_smooths(model = modGAM_95, series = AMF) +
  theme_bw() +
  geom_point(data=d3ta, aes(x=AMF, y=s.dw.ratioSNS, color=soil),  size=2) +
  annotate("text", x=50000, y=1.25, label= expression("AMF"), size = si) +
  annotate("text", x=50000, y=0.25, label= expression(paste(italic(r^2), "= 0.05")) , size = sz) +
  scale_color_manual(values=Kolor[2:4])+
  scale_y_continuous(breaks = c(0.5, 1)) +
    labs(y=expression(SDW[italic(Rel.)])) +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.title =element_text(size=axsz),
        axis.text = element_text(size=axtk),
          panel.grid = element_blank()) +
  coord_cartesian(ylim = c(0, 1.3)) 

# H soil
gp_H_FAKE_AMF <- ggplot(data=d3ta[d3ta$soil=="H",], aes(x=AMF, y=s.dw.ratioSNS)) + 
  theme_bw() +
  geom_point(aes(color=soil), size=2, alpha=0.6) +
  annotate("text", x=20000, y=1.25, label= expression("AMF"),  size = si, color="grey60") +
  scale_color_manual(values=Kolor[2]) +
  scale_y_continuous(breaks = c(0.5, 1)) +
  theme(legend.position = "none",
          axis.title=element_blank(),
          axis.text = element_text(color= "grey60", size=axtk),
          axis.ticks = element_line(color = "grey60"),
          panel.border = element_rect(color="grey60"),
          panel.grid = element_blank()) +
  coord_cartesian(ylim = c(0, 1.3)) # this is only a scatterplot for AMF. AMF is not included in the final model in H soil; but it is in the 3soil GAM model, so I display AMF-SDWrel scatterplot as shaded out version....


gp_H_Cros <- tidymv::plot_smooths(model = modGAM_H_100, series = Cros) +
  theme_bw() +
  geom_point(data=d3ta[d3ta$soil=="H",], aes(x=Cros, y=s.dw.ratioSNS, color=soil), size=2) +
  annotate("text", x=60, y=1.25, label= expression(italic("C. rosea")), size = si) +
  annotate("text", x=60, y=1.1, label= expression(paste(italic(r^2), "= 0.12")), size = sz) +
  scale_color_manual(values=Kolor[2]) +
  scale_y_continuous(breaks = c(0.5, 1)) +
   theme(legend.position = "none",
        axis.title= element_blank(),
        axis.text = element_text(size=axtk),
          panel.grid = element_blank()) +
  coord_cartesian(ylim = c(0, 1.3)) 

gp_H_FAKE_Rsol <- ggplot(data=d3ta[d3ta$soil=="H",], aes(x=Rsol, y=s.dw.ratioSNS)) +   # this is again just a scatterplot.
  theme_bw() +
  geom_point(aes(color=soil), size=2, alpha=0.6) +
  annotate("text", x=40, y=1.25, label= expression(italic("R. solani")), size = si, color="grey60") +
  scale_color_manual(values=Kolor[2]) +
  scale_y_continuous(breaks = c(0.5, 1)) +
  theme(legend.position = "none",
          axis.title=element_blank(),
          axis.text = element_text(color= "grey60", size=axtk),
          axis.ticks = element_line(color = "grey60"),
          panel.border = element_rect(color="grey60"),
          panel.grid = element_blank()) +
  coord_cartesian(ylim = c(0, 1.3)) 

gp_H_Fsol <- tidymv::plot_smooths(model = modGAM_H_100, series = Fsol) +
  theme_bw() +
  geom_point(data=d3ta[d3ta$soil=="H",], aes(x=Fsol, y=s.dw.ratioSNS, color=soil), size=2) +
  labs(y=expression(SDW[italic(Rel.)]),
       x=expression(paste("pg ", rct^-1))) +
  annotate("text", x=1250, y=1.25, label= expression(italic("F. solani")), size = si) +
  annotate("text", x=1250, y=1.1, label= expression(paste(italic(r^2), "= 0.29")) , size = sz) +
  scale_color_manual(values=Kolor[2]) +
  scale_y_continuous(breaks = c(0.5, 1)) +
  labs(x=expression("DNA amount")) +
  theme(legend.position = "none",
        axis.title.y= element_blank(),
        axis.title =element_text(size=axsz),
        axis.text = element_text(size=axtk),
          panel.grid = element_blank()) +
  coord_cartesian(ylim = c(0, 1.3)) 

# L soil
gp_L_AMF <- tidymv::plot_smooths(model = modGAM_L_100, series = AMF) +
  theme_bw() +
  geom_point(data=d3ta[d3ta$soil=="L",], aes(x=AMF, y=s.dw.ratioSNS, color=soil), size=2) +
  annotate("text", x=5000, y=1.25, label= expression("AMF"),  size = si) +
  annotate("text", x=5000, y=1.1, label= expression(paste(italic(r^2), "= 0.15")), size = sz) +
  scale_color_manual(values=Kolor[3]) +
  scale_y_continuous(breaks = c(0.5, 1)) +
  theme(legend.position = "none",
        axis.title=element_blank(),
        axis.text = element_text(size=axtk),
          panel.grid = element_blank()) +
  coord_cartesian(ylim = c(0, 1.3)) 

gp_L_FAKE_Cros <- ggplot(data=d3ta[d3ta$soil=="L",], aes(x=Cros, y=s.dw.ratioSNS)) +   # this is again just a scatterplot.
  theme_bw() +
  geom_point(aes(color=soil), size=2, alpha=0.6) +
  annotate("text", x=25, y=1.25, label= expression(italic("C. rosea")), size = si, color="grey60") +
  scale_color_manual(values=Kolor[3]) +
  scale_y_continuous(breaks = c(0.5, 1)) +
  theme(legend.position = "none",
          axis.title=element_blank(),
          axis.text = element_text(color= "grey60", size=axtk),
          axis.ticks = element_line(color = "grey60"),
          panel.border = element_rect(color="grey60"),
          panel.grid = element_blank()) +
  coord_cartesian(ylim = c(0, 1.3))

gp_L_FAKE_Rsol <- ggplot(data=d3ta[d3ta$soil=="L",], aes(x=Rsol, y=s.dw.ratioSNS)) +   # this is again just a scatterplot.
  theme_bw() +
  geom_point(aes(color=soil), size=2, alpha=0.6) +
  annotate("text", x=100, y=1.25, label= expression(italic("R. solani")), size = si, color="grey60") +
  scale_color_manual(values=Kolor[3]) +
  scale_y_continuous(breaks = c(0.5, 1)) +
  theme(legend.position = "none",
          axis.title=element_blank(),
          axis.text = element_text(color= "grey60", size=axtk),
          axis.ticks = element_line(color = "grey60"),
          panel.border = element_rect(color="grey60"),
          panel.grid = element_blank()) +
  coord_cartesian(ylim = c(0, 1.3)) 

gp_L_Fsol <- tidymv::plot_smooths(model = modGAM_L_100, series = Fsol) +
  theme_bw() +
  geom_point(data=d3ta[d3ta$soil=="L",], aes(x=Fsol, y=s.dw.ratioSNS, color=soil), size=2) +
  annotate("text", x=1500, y=1.25, label= expression(italic("F. solani")),  size = si) +
  annotate("text", x=1500, y=1.1, label= expression(paste(italic(r^2), "= 0.17")), size = sz) +
  scale_color_manual(values=Kolor[3]) +
  scale_y_continuous(breaks = c(0.5, 1)) +
  labs(x=expression("DNA amount")) +
  theme(legend.position = "none",
        axis.title.y=element_blank(),
        axis.title =element_text(size=axsz),
        axis.text = element_text(size=axtk),
          panel.grid = element_blank()) +
  coord_cartesian(ylim = c(0, 1.3)) 

# W soil
gp_W_AMF <- tidymv::plot_smooths(model = modGAM_W_100, series = AMF) +
  theme_bw() +
  geom_point(data=d3ta[d3ta$soil=="W",], aes(x=AMF, y=s.dw.ratioSNS, color=soil), size=2) +
  annotate("text", x=50000, y=1.25, label= expression("AMF"),  size = si) +
  annotate("text", x=50000, y=0.25, label= expression(paste(italic(r^2), "= 0.29")), size = sz) +
  scale_color_manual(values=Kolor[4]) +
  scale_y_continuous(breaks = c(0.5, 1)) +
  theme(legend.position = "none",
        axis.title =element_blank(),
        axis.text = element_text(size=axtk),
          panel.grid = element_blank()) +
  coord_cartesian(ylim = c(0, 1.3))

gp_W_FAKE_Cros <- ggplot(data=d3ta[d3ta$soil=="W",], aes(x=Cros, y=s.dw.ratioSNS)) +   # this is again just a scatterplot.
  theme_bw() +
  geom_point(aes(color=soil), size=2, alpha=0.6) +
  annotate("text", x=15, y=1.25, label= expression(italic("C. rosea")), size = si, color="grey60") +
  scale_color_manual(values=Kolor[4]) +
  scale_y_continuous(breaks = c(0.5, 1)) +
  theme(legend.position = "none",
          axis.title=element_blank(),
          axis.text = element_text(color= "grey60", size=axtk),
          axis.ticks = element_line(color = "grey60"),
          panel.border = element_rect(color="grey60"),
          panel.grid = element_blank()) +
  coord_cartesian(ylim = c(0, 1.3))

gp_W_Rsol <- tidymv::plot_smooths(model = modGAM_W_100, series = Rsol) +
  theme_bw() +
  geom_point(data=d3ta[d3ta$soil=="W",], aes(x=Rsol, y=s.dw.ratioSNS, color=soil), size=2) +
  annotate("text", x=150, y=1.25, label= expression(italic("R. solani")), size = si) +
  annotate("text", x=150, y=1.1, label= expression(paste(italic(r^2), "= 0.18")), size = sz) +
  scale_color_manual(values=Kolor[4]) +
  scale_y_continuous(breaks = c(0.5, 1)) +
  theme(legend.position = "none",
        axis.title=element_blank(),
        axis.text = element_text(size=axtk),
          panel.grid = element_blank()) +
  coord_cartesian(ylim = c(0, 1.3)) 

gp_W_Fsol <- tidymv::plot_smooths(model = modGAM_W_100, series = Fsol) +
  theme_bw() +
  geom_point(data=d3ta[d3ta$soil=="W",], aes(x=Fsol, y=s.dw.ratioSNS, color=soil), size=2) +
  annotate("text", x=750, y=1.25, label= expression(italic("F. solani")), size = si) +
  annotate("text", x=750, y=0.25, label= expression(paste(italic(r^2), "= 0.16")), size = sz) +
  scale_color_manual(values=Kolor[4]) +
  scale_y_continuous(breaks = c(0.5, 1)) +
  theme(legend.position = "none",
        axis.title.y =element_blank(),
        axis.title =element_text(size=axsz),
        axis.text = element_text(size=axtk),
          panel.grid = element_blank()) +
  coord_cartesian(ylim = c(0, 1.3)) +
  xlab("DNA amount")




### Titles
szo <- 4

title_3 <- ggplot(mtcars, aes(wt, mpg)) +
  annotate("text", x=2, y=2,  hjust = 0, label= "3 infested soils", size = szo) +
  theme_void() +
  xlim(0,11) +
  ylim(0,4)

title_H <- ggplot(mtcars, aes(wt, mpg)) +
  annotate("text", x=2, y=2,  hjust = 0, label= "Kirchlindach", size = szo) +
  theme_void() +
  xlim(0,11) +
  ylim(0,4)

title_L <- ggplot(mtcars, aes(wt, mpg)) +
  annotate("text", x=2, y=2,  hjust = 0, label= "Puch", size = szo) +
  theme_void() +
  xlim(0,11) +
  ylim(0,4)

title_W <- ggplot(mtcars, aes(wt, mpg)) +
  annotate("text", x=2, y=2,  hjust = 0, label= "Neu-Eichenberg", size = szo) +
  theme_void() +
  xlim(0,11) +
  ylim(0,4)


# Put all together
library(cowplot)
### use NULL for empty plots. The alphabetical order of all fungi is: AMF, Apha, Cros, Fave, Foxy, Fredo, Fsol, Phoma, Pult, Rsol


titles <- plot_grid(title_3, title_H, title_L, title_W, ncol = 4)
                
bigplot <- plot_grid(gp_3_AMF,  gp_H_FAKE_AMF,  gp_L_AMF,       gp_W_AMF,
                     gp_3_Cros, gp_H_Cros,      gp_L_FAKE_Cros, gp_W_FAKE_Cros,
                     gp_3_Rsol, gp_H_FAKE_Rsol, gp_L_FAKE_Rsol, gp_W_Rsol,
                     gp_3_Fsol, gp_H_Fsol,      gp_L_Fsol,      gp_W_Fsol,
          rel_widths = c(1.1,1,1,1,
                         1.1,1,1,1,
                         1.1,1,1,1,
                         1.1,1,1,1),
          rel_heights = c(0.9,0.9,0.9,0.9,
                          0.9,0.9,0.9,0.9,
                          0.9,0.9,0.9,0.9,
                          1.1,1.1,1.1,1.1),
          ncol = 4) 

bibigplot <- plot_grid(titles, bigplot, ncol = 1, rel_heights = c(1,10))


ggsave("GAMplots_3HLW_r2.tiff", bibigplot, units="cm", width=18, height=18, dpi=300, device = "tiff")
  
```



### Network analysis: Graphical model with ecoCopul (Popovic et al.)

This analysis is made in two parts: a) formulate ultivariate community data model with mvabund(). I will fit a negative binomial model, and b) the actual graphical model based on that former regression model

0) data prep
```{r ecoCopula graphical model - data prep}
Topf18.FUNGI.MEANS.3sick <- Topf18.FUNGI.MEANS.NONA[Topf18.FUNGI.MEANS.NONA$soil!="F",]
Topf18.FUNGI.MEANS.H <-     Topf18.FUNGI.MEANS.NONA[Topf18.FUNGI.MEANS.NONA$soil=="H",]
Topf18.FUNGI.MEANS.L <-     Topf18.FUNGI.MEANS.NONA[Topf18.FUNGI.MEANS.NONA$soil=="L",]
Topf18.FUNGI.MEANS.W <-     Topf18.FUNGI.MEANS.NONA[Topf18.FUNGI.MEANS.NONA$soil=="W",]

T18.3sick.fungiquants <- round(mvabund::mvabund(Topf18.FUNGI.MEANS.3sick[, c(7:16)]))
T18.H.fungiquants <- round(mvabund::mvabund(Topf18.FUNGI.MEANS.H[, c(7:16)]))
T18.L.fungiquants <- round(mvabund::mvabund(Topf18.FUNGI.MEANS.L[, c(7:16)]))

T18.W.fungiquants <- round(mvabund::mvabund(Topf18.FUNGI.MEANS.W[, c(7:16)]))

```


So: because I have some fungi that do not have 0's and the NB model generally is fine I have good reasons to continue with a NB modelling in the mvabund... I also tried linera model (also mit log transformation; their residual plots all do not look good)


```{r ecoCopula graphical model - model formulation}
manyglm_3sick <- manyglm(T18.3sick.fungiquants ~ Topf18.FUNGI.MEANS.3sick$soil, family="negative_binomial") # here the abundances are modelled wit hthe factor "soil""
ccgr3sick <- cgr(manyglm_3sick)

manyglm_H <- manyglm(T18.H.fungiquants ~ 1, family="negative_binomial") #within the soils the abundances are modelled without factors
ccgrH <- cgr(manyglm_H)

manyglm_L <- manyglm(T18.L.fungiquants ~ 1, family="negative_binomial")
ccgrL <- cgr(manyglm_L)

manyglm_W <- manyglm(T18.W.fungiquants ~ 1, family="negative_binomial")
ccgrW <- cgr(manyglm_W)

```

```{r ecoCopula graphical model - plot networks}
# prep
library(ggraph)
labels <- c("A. eut.", "F. ave.", "F. oxy.", 'F. red.', 'F. sol.', 'D. pin.', 'P. ult.', 'R. sol.', "AMF", 'C. ros.')
labelsize = 2.4
mrgs=c(2,4,2,4)
colbad <- "white"
colgood <- "#999999"

# 4 plots
net3 <- ccgr3sick$best_graph$igraph_out %>%
  ggraph("fr") + 
  geom_edge_fan0(aes(linetype = cut(partcor, c(-1,0,1)), width=partcor)) +
  scale_edge_width(range = c(0.3, 3)) +
  scale_edge_linetype_manual(values=c("longdash", "solid" ), name="sign.") +
  geom_node_label(aes(label=labels), fontface=c(rep(3,8), 2, 4), size=labelsize) +
  theme_void() +
  ggtitle("3 infested soils") +
  theme(plot.title = element_text(size=10, vjust = 0.7), plot.margin = unit(mrgs, "mm")) +
  coord_cartesian(clip = "off")

netH <- ccgrH$best_graph$igraph_out %>%
  ggraph("fr") +
  geom_edge_fan0(aes(linetype = cut(partcor, c(-1,0,1)), width=partcor), color=Kolor[2]) +
  scale_edge_width(range = c(0.3, 3)) +
  scale_edge_linetype_manual(values=c("longdash", "solid" )) +
  geom_node_label(aes(label=labels), fontface=c(rep(3,8), 2, 4), size=labelsize, color=Kolor[2]) +
  theme_void() +
  ggtitle("Kirchlindach") +
  theme(plot.title = element_text(size=10, vjust = 0.7), legend.position = 'none', plot.margin = unit(mrgs, "mm")) +
  coord_cartesian(clip = "off")

netL <- ccgrL$best_graph$igraph_out %>%
  ggraph("fr") + 
  geom_edge_fan0(aes(linetype = cut(partcor, c(-1,0,1)), width=partcor), color=Kolor[3]) +
  scale_edge_width(range = c(0.3, 3)) +
  scale_edge_linetype_manual(values=c("longdash", "solid" )) +
  geom_node_label(aes(label=labels), fontface=c(rep(3,8), 2, 4), size=labelsize, color=Kolor[3]) +
  theme_void() +
  ggtitle("Puch") +
  theme(plot.title = element_text(size=10, vjust = 0.7), legend.position = 'none', plot.margin = unit(mrgs, "mm")) +
  coord_cartesian(clip = "off")

netW <- ccgrW$best_graph$igraph_out %>% 
  ggraph("fr") + 
  geom_edge_fan0(aes(linetype = cut(partcor, c(-1,0,1)), width=partcor), color=Kolor[4]) +
  scale_edge_width(range = c(0.3, 3), name="Partial\ncorr.") +
  scale_edge_linetype_manual(values=c("longdash", "solid" )) +
  geom_node_label(aes(label=labels), fontface=c(rep(3,8), 2, 4), size=labelsize, color=Kolor[4]) +
  theme_void() +
  theme(legend.position = 'none', plot.margin = unit(mrgs, "mm"), plot.title =   element_text(size=10, vjust = 0.7)) +
  ggtitle("Neu-Eichenberg") +
  coord_cartesian(clip = "off") 


# put together
thelegend <- ggpubr::get_legend(net3 + theme(legend.box.margin = margin(3, 0, 0, 5)))

plot4s <- cowplot::plot_grid(net3 + theme(legend.position="none"),
                             netH, netL, netW,
                   nrow = 2)

theplot <- cowplot::plot_grid(plot4s, thelegend, rel_widths = c(10,1))

#ggsave("ecoCopula_4graphs_raw.tiff", theplot, units="cm", width=18, height=10, dpi=300, compression = 'lzw')

```






